Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-07-28 pirouette_supplement.tex
Address more points of confusion

To assess if the results of the inference are meaningful one important 
parameter is the Effective Sample Size (ESS). This quantity describes how 
many independent trees are sampled from the posterior distributions. 
For reliable results it is good practice to have at 
least $ESS = 200$ (see 
\begin{sloppypar}
  \url{https://beast.community/ess_tutorial}).
\end{sloppypar}
In the following we present the ESS for the posterior distributions of 
the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}: 
"true" pipeline with generative model (Table \ref{tab:esses_gen}), 
"true" pipeline with best candidate model (Table \ref{tab:esses_best}), 
"twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), 
"twin" pipeline with best candidate model (Table \ref{tab:esses_twin_best}). 
We also report the marginal likelihood (or evidence) data for model selection 
performed both in the "true" (Table \ref{tab:evidence_true}) and 
"twin" pipeline (Table \ref{tab:evidence_twin}).

@@

@@

@@

@@

@@

>>>>

@@

To assess if the results of the inference are meaningful one important 
parameter is the Effective Sample Size (ESS). This quantity describes how 
many independent trees are sampled from the posterior distributions. 
For reliable results it is good practice to have at 
least $ESS = 200$ (see 
\begin{sloppypar}
  \url{https://beast.community/ess_tutorial}).
\end{sloppypar}
In the following we present the ESS for the posterior distributions of 
the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}.

The ESSes of the 'true' pipeline for the generative model
are shown in Table \ref{tab:esses_gen}.
From the estimated parameters, one can deduce that
the JC69 nucleotide substitution model was used (no
estimated parameter needed), a strict clock model was 
used (again, no parameter needed to be estimated)
and a Yule tree prior is used ('Yule model' and 'birthRate' are estimated).
Note that although the actual true tree is created by a DD process, 
the default and standard Yule tree model is used as the closest
standard tree model.

@@

The ESSes of the 'twin' pipeline for the generative model
are shown in Table \ref{tab:esses_twin_gen}.
Note that the generative inference model is 
re-used (which assumes a Yule tree model) in the inference, 
where the twin tree is actually created using a BD process,
which is the default.

The ESSes of the 'true' pipeline for the best candidate model
are shown in Table \ref{tab:esses_best}. 
From the names of the estimated parameters, it is clear that
the best candidate model has had a TN93 nucleotide substitution 
model (due to 'kappa1', 'kappa2', 'freqParameter.1', 
'freqParameter.2', 'freqParameter.3', 'freqParameter.4')
a strict clock (no parameter needed to be estimated) and a BD 
model ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').

@@

The ESSes of the 'twin' pipeline for the best candidate model
are shown in Table \ref{tab:esses_twin_best}.

From the names of the estimated parameters, it is clear that
the best candidate model for the twin tree
is JC69 nucleotide substitution 
model (no parameter needed to be estimated),
a strict clock (no parameter needed to be estimated) and a BD 
model ('BirthDate', 'BDBirthRate' and 'BDDeathRate').
Note that this matches the actual process of how the twin tree
and twin alignment are generated.

@@

The marginal likelihood (or evidence) data for the model comparison
performed in the 'true' pipeline is shown in Table \ref{tab:evidence_true}.
The best (that is, the one with the highest model weight)
candidate model assumes a TN93 nucleotide substitution model,
a strict clock and a BD model.

@@

The marginal likelihood (or evidence) data for model comparison
performed in the 'twin' pipeline is shown in Table \ref{tab:evidence_twin}.
The best (that is, the one with the highest model weight)
candidate model assumes a JC69 nucleotide substitution model,
a strict clock and a BD model.
Note that this matches the actual process of how the twin tree
and twin alignment are generated.
